---
title: "Good Coding Practices"
author: "Andrew Do"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sol = TRUE
```

## Short Exercises
Rewrite the following according to our class style guide:

1.
```{r, eval = FALSE}
x < - c( NA,1,-2,3,-4,5 )
  y<-3
  y<-x * - 1
  y[ y>0 ]
```

```{r, echo = sol}

x <- c(NA,1,-2,3,-4,5)
  y <- 3
  y <- x*(-1)
  y[y > 0]
  
```

2. 
```{r}
mean <- function(x) {
sum (x)/length( x )
}
```

```{r, echo = sol}

mean <- function(x) {
  
#Caclulates the average of the function 
  #Arguments: x = the vector whose values are to be calculated 
  #Returns the average of the input x

sum(x)/length(x)
  
}

```

3. 
```{r}
library( ggplot2 )
second_version_of_mpg <- mpg[ mpg$cyl == 6,]
second_version_of_mpg$class <- as.character(second_version_of_mpg$class)
ggplot(data=second_version_of_mpg,aes(cty,hwy,color=class))+geom_point()+geom_smooth( method = "lm", se=FALSE ) + labs (title="City vs. Highway MPG" )
```

```{r, echo = sol}

library(ggplot2)

second_version_of_mpg <- mpg[mpg$cyl == 6, ]
second_version_of_mpg$class <- as.character(second_version_of_mpg$class)

ggplot(data=second_version_of_mpg, aes(cty, hwy, color=class)) + 
  geom_point() + 
  geom_smooth(method = "lm", 
              se = FALSE) + 
  labs(title = "City vs. Highway MPG")

```

## Longer Exercises

Fix the following code with the following goals:

* Readability - an untrained person should be able to guess what's going on
    + Use meaningful names
    + Use whitespace to make the code flow
    + Provide documentation
    + Use functions from the packages we've covered (they're designed to be readable!)
  
* Compactness - do not be more verbose than you have to be
    + Think of ways to vectorize the function
    + Avoid copy-paste
    + Use the `lapply` family
    + Modularize your code (small functions instead of massive ones)
  
* Usable as intended
    + If necessary, include error messages
    + Don't force the user to rely on implicit coercion
    + Output the intentional data type

### Exercise 1
Infer what the function `d` is doing and rewrite it according to our class style guide.  Include documentation so that future users won't have to decipher the code.
```{r}

d <- function(f, x, y, s) {
  stuff<-seq(x,y,by=s)
  stuff2<-c()
  for(i in 1:length(stuff)){
    
    thing<-f(stuff[i]); stuff2[i]<-thing
    
    }
  m<-max(stuff2)
  mm<-stuff[stuff2==m]
  mm<-min(mm)
  return(list(m,mm))
}

# Example usage
g <- function(x) -(x ^ 2)
d(g, -1, 1, .1)
```

```{r, echo = sol}

max_min <- function(f, x, y, s) {
  
  #This function is meant to receieve four arguements  
  
  counter <- seq(x, y, by = s)
  emptyvec <- c()
  
  #Create an empty vector for the arguments, and list out a sequence using the arguments 
  
  for(i in 1:length(counter)) {
    
    input <- f(counter[i])
    emptyvec[i] <- input
    
  }
  
  maxvalue <- max(emptyvec)
  minvalue <- min(counter[emptyvec == maxvalue])

  
  return(list(maxvalue, minvalue))
#returns max, min of function 
  }

# Example usage
samplefunc <- function(x) - (x^2)
max_min(samplefunc, -1, 1, .1)

```


### Exercise 2
Figure out what the `ss` function is doing, rewrite it so it uses more expressive functions according to our style guide.  Provide documentation.
```{r}
ss <- function(x) 
{x<-x[!is.na(x)]
a<-x[1];for(i in 2:length(x)) a[i]<-(a[i-1]+x[i])
a<-a/(1:length(x))
foo<-x
cc<-0
for(i in 1:length(foo)){x<-foo[1:i]
c<-sort(x);
if(length(x)%%2==0){cc[i]<-(c[floor((length(c)+1)/2)] + c[ceiling((length(c)+1)/2)])/2} else{ cc[i]<-c[(length(c)+1)/2]}}
return(list(a,cc))
}

# Example usage
ss(c(5,10,15,20,5,10,15,20))
ss(1:10)
ss(c(1,2,6,2,4,5))
```

```{r, echo = sol}

decimals <- function(x) {
  
# This function has one input/argument of x

detector <- x[!is.na(x)]
#If there are NA values, then show as False
output <- detector[1]

#inspect the length of the vector 
for(i in 2:length(x)) { 
  
  output[i] <- (output[i-1] + detector[i])
  output <- output/(1:length(x))

  }

record <- detector
counter <- 0

#initialize a counter variable

for(i in 1:length(record)) {
  
  detector <- record[1:i]
  organize <- sort(x)
  
}

#Decide whether a certain output is needed versus another

if(length(x) %%2 == 0) {
  
  counter[i] <- (organize[floor((length(organize) + 1)/2)] + organize[ceiling((length(organize) + 1)/2)])/2
  
} 

#If the first statement proves false, then move onto this else statement

else{ 
  
  counter[i] <- organize[(length(organize)+1)/2]
  
}

return(list(output,counter))
#return the output and counter from the previous calls
}


# Example usage
decimals(c(5,10,15,20,5,10,15,20))
decimals(1:10)
decimals(c(1,2,6,2,4,5))
```

### Exercise 3
Rewrite the following script in a more compact form using packages we've covered in this class.  It shouldn't take more than 7 function calls---3 if you took the time to read the `dplyr` vignette!
```{r}

d<-iris
dd<-unique(d[,5])
d1<-d[which(d[,5]==dd[1]),]
d2<-d[which(d[,5]==dd[2]),]
d3<-d[which(d[,5]==dd[3]),]
d1[,1] <- mean(d1[,1])
d1[,2] <- mean(d1[,2])
d1[,3] <- mean(d1[,3])
d1[,4] <- mean(d1[,4])
d2[,1] <- mean(d2[,1])
d2[,2] <- mean(d2[,2])
d2[,3] <- mean(d2[,3])
d2[,4] <- mean(d2[,4])
d3[,1] <- mean(d3[,1])
d3[,2] <- mean(d3[,2])
d3[,3] <- mean(d3[,3])
d3[,4] <- mean(d3[,4])
ddd<-rbind(d1[1,],d2[1,],d3[1,])
rownames(ddd) <- NULL



```

```{r, echo = sol}


library(dplyr)
#install dplyr packages

data = group_by(iris, Species)

#group by the different species 

output = summarize_each(data, funs(mean))

#output the mean of each column, seperated by species

output
```

## Remote Computing Practice
Upload the `.rmd` file to `Documents/lab19/` on your server account and knit it using the command line. Submit the `.html` file produced by knitting on the server.  Assuming you're in the folder with the markdown file on the server, the command is:

```{sh, eval = FALSE}
Rscript -e "rmarkdown::render('lab19.Rmd')"
```


```{r, echo = sol}
sessionInfo()
```

