---
title: "hw8"
author: "Due: 11:59pm Saturday 8/6/2016.  Instructors stop answering questions at 5pm."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(dplyr)
library(ggplot2)
library(readr)
library(RSQLite)
library(purrr)
library(ggmap)
library(tidyr)
library(stringr)
```

### purrr

In this exercise, you'll be using functional programming ideas to simulate the central limit theorem on coin flips.  Using neither loops nor the apply functions, write a simulation pipeline that:

* Generates 1000 samples of each 1, 2, ..., 99, 100 fair coin flips.  So for example, you would simulate flipping 1 coins 1000 times, 2 coins 1000 times, etc.
* For each of the 1, 2, ..., 100 flips, produces a histogram of the proportion of heads.  To be explicit: you will create a histogram for the 1000 proportions calculated from the 1 coin flips.  Then you'll do the same for 2 coin flips, 3, etc.

In your plots
* The x-axis should be limited to 0-1
* The y-axis should be limited to 0-15


```{r, eval = FALSE}

flip = function(x) {
  samp = sample(c(1.00,0.00), 
             size = x, 
             replace = TRUE,
             prob = c(0.5,0.5))
  proportion = sum(samp)/x
  return(proportion)
}

test = 1:100
results = test %>% 
  map(function(x) rerun(1000, flip(x)))


plot = function(b) {
  coin = as.data.frame(results[[b]]) %>%
    gather(key = count,
           value = proportion) %>% 
    mutate(count = c(1:1000))
  
  ggplot(coin) +
    aes(x = proportion, y = ..density..) +
    geom_histogram(bins = 30) +
    labs(x = "Proportion of heads",
         y = "Density",
         title = "Distribution of Heads") +
    scale_y_continuous(limits = c(0,15)) +
    scale_x_discrete(limits = c(0,1))
}

map(test, plot)
```

Now that you have all the static images, you could "animate" the progression by turning the sequence of images into a gif.  This can be achieved with the `animation` package in R or the standalone `ImageMagick` utility.  This is just an FYI, we will not be looking for a GIF.  Also, Please do NOT upload the images to GitHub nor bCourses.  We'll just be checking your code.

## SQL

These exercises are meant to help you wrap your head around SQL.  Answer the questions below using SQL queries.  You can check your answer using `dplyr` methods, but your grade will be based on the SQL commands.  The `nycflights13` data set has been put into a SQLlite database for these problems.  For documentation about the variable names, see

```{r, eval = FALSE}
help(package = "nycflights13")

answers = read_rds("hw8_sql_results (1).rds")
answers
```

```{r}
flights <- src_sqlite("nycflights13.mysqlite3")
```


1. Find the full name of each airline that flew to Dallas Fort-Worth (DFW).
```{r}



airlinesnames = flights %>% tbl(
  
  sql("SELECT DISTINCT 
nyairlines.name
FROM 
nyflights LEFT JOIN nyairlines 
ON nyflights.carrier = nyairlines.carrier
WHERE dest = 'DFW'
               ")) %>% collect(n = Inf)



all.equal(answers[[1]], airlinesnames)

```

2. Make a table containing the tail number, year of manufacture, model, number of engines, and number of seats of the planes flown by United Airlines (UA).  Sort the results by year manufactured.

```{r}

newtable = flights %>% tbl(
sql("SELECT DISTINCT
nyplanes.year, 
nyplanes.tailnum, 
nyplanes.model, 
nyplanes.engines,
nyplanes.seats
FROM 
nyflights INNER JOIN nyplanes
ON nyplanes.tailnum = nyflights.tailnum 
WHERE 
carrier = 'UA'
ORDER BY
nyplanes.year")) %>% collect (n = Inf)

newtable
all.equal(answers[[2]], newtable)
```

3. Make a table with the following information: full airport name in a column called `Airport` and number of flights from New York City to those airports in 2013 in a column called `NumberOfFlights`.  Sort the results in descending order by flight counts.
```{r}

airportinfo = flights %>%  tbl(
  sql( "SELECT 
       nyairports.name AS Airport, 
       COUNT(nyairports.name) AS NumberOfFlights
      FROM 
    nyairports LEFT JOIN nyflights
       ON nyairports.faa = nyflights.dest
    WHERE 
    year = 2013
    GROUP BY
    Airport
    ORDER BY
    count(nyairports.name) DESC
    ")) %>% 
  collect(n = Inf)

airportinfo 

all.equal(answers[[3]], airportinfo)


```

4. Find the wind speed during the hour of the sceduled departure time of every flight that had a departure delay of more than 30 minutes.  For example, if a flight was scheduled to leave at 5:59am, you would check the wind speed for 5:00am.  In practice, you would of course be more careful than this.
```{r}

wind = flights %>% tbl(
   sql("SELECT
    nyweather.wind_speed,
    nyflights.time_hour,
    nyflights.hour
    FROM 
    nyweather LEFT JOIN nyflights
    ON 
    nyflights.origin = nyweather.origin AND 
    nyflights.time_hour = nyweather.time_hour
    WHERE 
    nyflights.dep_delay > 30")) %>% collect(n = Inf)

wind
all.equal(answers[[4]], wind)

```