---
title: "lab25"
date: "July 26, 2016"
output: html_document
---

Write a function called `moveCars()` that moves the system one full step; that is, the function should move the red cars east once and the blue cars north once, with red cars going first.

Your function should take the following inputs:
`m`: a matrix

Your function should return a list containing two elements:
`m`: an updated matrix
`not_gridlocked`: a logical value that should be TRUE if the system changed and FALSE if not

```{r, eval = FALSE}


moveCars = function(m) {
  n = 1
  height = dim(m)[1]
  width = dim(m)[2]
  
  
#### This section focuses on moving the blue cars 
  
  
  if(height != 1){
    blue = output
    if(width == 1){
      zero_blocked = (matrix(c (output[height,]
                             output[1:(height-1),] ))==0)
    } 
    
  }
  zero = !identical(m,output)
  m = output
  return(list(m,
              zero))
}
  
  
## This section focuses on moving the red cars 


  if(width != 1){
    red = m * (m==1)
    } else{
      none = (cbind (m[,2:width],m[,1])==0)
    }
    move = none*red 
  
```


Check:
```{r, eval = FALSE}
m <- matrix(c(1,0,0,0,2,1,2,0,0,2,0,1,1,0,1,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(0,1,0,0,2,1,2,0,0,2,1,1,0,1,0,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T), TRUE))

m <-  matrix(c(0,1,0,2,1), nrow = 1, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(1,0,1,2,0), nrow = 1, ncol = 5, byrow = T), TRUE))

m <- matrix(c(2,0,1,2,0), nrow = 5, ncol = 1, byrow = T)
identical(moveCars(m), list(matrix(c(0,0,1,2,2), nrow = 5, ncol = 1, byrow = T), TRUE))
```

Now write a function called `initGrid()` that randomizes the initial system.

Your function should take the following inputs:
`r`: number of rows in the matrix
`c`: number of columns in the matrix
`d`: density of the grid 

Your function should return the following:
`m`: an rxc matrix with entries 0 (no cars), 1 (red cars), or 2 (blue cars)

```{r}


initGrid = function(r, c, d)
  
{
  
  hi = sample(c(0,1,2), r*c, prob = c(1 - d, d/2, d/2), replace = TRUE)
  
  matrix(hi, nrow = r, ncol = c)
  
}



initGrid(4,5, 1)

```






