---
title: "lab4"
author: "Andrew Do"
date: "June 13, 2016"
output: html_document
---


```{r setup, include=FALSE}
library(stringr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)

```

## Getting Help for Packages

You can see what an entire package has to offer with
```{r}
help(package = "stringr")
```


## Counting number of vowels

Consider the following vector `letrs` which contains various letters:
```{r}
# Garbage string
letrs <- "gfngenodgsyctyviqyyigeinyngbkwjdjqkrrollhpxdkfldlytympmcgmjynihusebtq"
```

If you were to count the number of vowels in `letrs` you would get the following counts:

- a: 0
- e: 3
- i: 4
- o: 2
- u: 1

This exercise will guide you through how to arrive at this frequency table.

```{r}
# Create a character vector of vowels.  Give it an appropriate name.

vowels = c("a", "e", "i", "o", "u")

# Use str_count to get the counts.  Assign this an appropriate name.

number = str_count(letrs, vowels)

# Give each element of the vector a name using the names function.

names(number) = vowels
 

number

# The result should look like:

# a e i o u

# 0 3 4 2 1
```

## Exploring new functions: str_pad
Write a short description of what each of the following calls of `str_pad` is doing.  The last three examples were found at the bottom of the function's documentation.

1. In this call of `str_pad`, This call is creating an output of length 4, and going from left to right, outputs three x's as "padding" before outputting the string "a". The padding is added to the left. 
```{r}
str_pad("a", width = 4, side = "left", pad = "x")
```

2. In this call of `str_pad`, This call is creating an output of length 3, and going from left to right has blank spaces for two spaces and ending with "a". This is similar to above, except no padding with x's and a shorter length. 
```{r}
str_pad("a", width = 3, side = "left")
```

3. In this call of `str_pad`, This call outputs a character vector, each component with a length of 10. The string is put to the right side, with space as padding for the other places. For "a", there are 9 empty spaces, "abc" has 7, and "abcdef" has 4. 
```{r}
str_pad(c("a", "abc", "abcdef"), 10)
```

4. In this call of `str_pad`, This call outputs "a" with different lengths of 5, 10, 20. Since the padding isn't specified, the a is immediately placed at the very right, with empty space filling up with lengths of 4, 9, and 19. 
```{r}
str_pad("a", c(5, 10, 20))
```

5. In this call of `str_pad`, The call outputs with "a", with three different sets of output with length 10. the padding is specified by a vector with three components, so there is padding with -, _, and empty space that fills up 9 spaces before "a". 
```{r}
str_pad("a", 10, pad = c("-", "_", " "))
```

## Regular Expressions

We read in the data from the csv file and extract the names column
```{r}
flags <- read_csv("flags.csv")
countries <- flags$name
```

You may have noticed that in the `countries` vector, words in a country's name are separated by `-`.  Use `str_replace_all` to replace the hyphens with spaces.  For example `"Vatican-City"` should be replaced with `"Vatican City"`

```{r}
# Your code here

str_replace_all(countries, "-", " ")
```

Use `str_detect` to see if there are any countries with the world "island" in them, ignoring case.  You may also want to look up how to use `ignore.case`.

```{r}
# Your code here

x = any(str_detect(countries, ignore.case("island")))
x




```

Use `str_subset` to find all countries that begin with a capital `A`.
```{r}
# Your code here

str_subset (countries, "^A")

```

Find all countries that end with `stan`
```{r}
# Your code here

str_subset (countries, "stan$")

```

Use `str_replace_all` to replace the `$`'s in `baws` with `s`'s.  Recall that the `$` is a metacharacter.
```{r}
baws <- "Bo$$"
str_replace_all(baws, "\\$", "s")
```

Use a character class (denoted by square brackets) to find all countries beginning with a vowel and ending with a consonant.
```{r}
# Fill in

a = "^[aeiouAEIOU]. + [^aeiou].$"


str_subset(countries, a) 


```




