---
title: "lab3"
author: "Andrew Do"
date: "June 8, 2016"
output: html_document
---

getwd()

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing the Data
Run install the package `readr` by running `install.packages("readr")` in the console.

The code below reads in the `tents.csv` comma-separated file.  The `col_types` argument is telling `R` what the data type is in each column.  See `?read_csv` if you're curious.
```{r}
library(readr)
setwd("~/Documents/stat133-su2016/03-lesson/lab")
tents <- read_csv("tents.csv", col_types = "ccnniccc")
```

### Questions
Answer the following questions programatically.  Some questions ask for you to explain what you see---you may answer this as text outside a code block.

1. Run `str()` on `tents`.  What information does it tell you? This is a useful function to know.
```{r} 

str(tents)

```
It tells me the name, brand, price, weight, height, bestuse, seasons, and capacity of different tents

2. What class is `tents`?

```{r}
class(tents)
```

Tents is a data frame. 

3. How many rows are in `tents`?

90 rows. 

```{r}
nrow(tents)
```


4. How many columns are in `tents`?

```{r}
ncol(tents)
```

8 columns 

5. What are the names of the columns?

name, brand, price, weight, height, bestuse, seasons, capacity

6. Display the first 5 rows.

```{r}
head(tents, 5)
# or
tents[seq(1,5),1:ncol(tents)] 
```


7. Display the last 4 rows.
```{r}
tail(tents,4)
# or
tents[seq(nrow(tents)-4, nrow(tents)),1:ncol(tents)]
```

8. Find out how to use `quantile` to find the interquartile range (75th percentile - 50th percentile) of the weights of the tents.

```{r}

quantile (tents[["weight"]], 0.75) - quantile(tents[["weight"]], 0.5)

```

IQR = 2350.625

9. Find the standard deviation of the heights of the `tents`.  Does `R` use `n` or `n-1` in the denominator when calculating standard deviation?

```{r}
sd (tents[,5])
```
This uses n -1 becuase of Bessel's Correction. 

10. The weight is given in grams.  Add a new column to the data frame called `weight_lbs` that contains the weight of the tents in pounds.  Remember to overwrite `tents` with your update.

```{r}


tents["weight_lbs"] = tents[["weight"]] * (1/453.592)

tents

```



11. The height is given in centimeters.  Add a new variable `height_in` to the data frame that contains the height of the tents in inches. Remember to overwrite `tents` with your update.

 
```{r}
tents["height_in"]= tents [["height"]] * 0.39370

tents
```


12. How many tents cost more than $400?

22 tents

```{r}
sum (tents [,3] > 400)
```

13. Subset the data frame so that it only contains tents that are valued $300 or less.  Assign the subsetted data frame the name `so_cheap_tents`.

tents 

```{r}

so_cheap_tents = tents[tents [["price"]] <= 300,1:ncol(tents)]

so_cheap_tents 
```

14. Of the tents that are more than $300, what percent are `big-agnes` tents?

```{r}



expensive_tents = tents [tents[["price"]] > 300, 1:ncol(tents)]

expensive_tents

total = sum(tents[["price"]] > 300)
bigA = sum(expensive_tents[["brand"]]== "big-agnes")

(bigA/total) * 100 

```


15. Select the `bestuse` column of `tents` in four different ways: using `$`, `[[]]`, position number and `[]`, and variable name and `[]`. 

```{r}

tents[["bestuse"]]

tents$bestuse

tents [1:nrow(tents), 6]

tents ["bestuse"]

```


16. What is the name of the tend with the best height-to-weight ratio?

```{r}

weight = tents[["weight"]]
height = tents [["height"]]
ratio = height/weight

tents [which.max(ratio),"name"]

```


17. Run `table` with two arguments, `seasons` and `capacity` (this should be one table).  What did it do?
```{r}

table(tents$seasons, tents$capacity)


```

This made a table with the number of times tents had a value with each cell.

18. What type of tent (bestuse) is the most common? Your answer should return only the use and not the count.

```{r}

unique(tents$bestuse)

##Backpacking, Mountaineeing, Carcamping

BP = nrow(tents[tents$bestuse == "Backpacking",])

M = nrow(tents[tents$bestuse == "Mountaineering",])

C = nrow(tents[tents$bestuse == "Carcamping",])

x = c(BP, M, C)
names = c("Backpacking", "Mountaineering", "Carcamping")

df = data.frame(x,names)

final = which.max(x)

df[final, "names"]

```



19. Create a data frame called `kelty` with only data on kelty tents.  Do the same for `big-agnes`.  Then, combine the two data frames.


```{r}

kelty = tents [tents[["brand"]] == "kelty", 1:ncol(tents)] 
kelty

big_agnes = tents [tents[["brand"]] == "big-agnes", 1:ncol(tents)]
big_agnes

total = rbind(kelty, big_agnes)
total

```
